Explore spatial allocation

### 2021-03-23 marine_mapping_archive_v1/crustaceans/crustacean_farms/STEP3_rasterize_production.Rmd
Finished reorganizing data

---
title: "Rasterizing prodution per each region"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

farms <- read.csv(here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon.csv"))

@@

# CYP:XAD
# ESP:PRT
# GRC:ALB
# GRC:TUR
# KOR:PRK
# HKG:CHN

mapview(filter(farms_sf, check=="CYP:XAD"))
mapview(filter(farms_sf, check=="ESP:PRT"))
mapview(filter(farms_sf, check =="GRC:ALB"))
mapview(filter(farms_sf, check == "GRC:TUR"))
mapview(filter(farms_sf, check == "KOR:PRK"))

# some annoying mismatches due to raster resolution. Fix these

unique(farms_sf$NAME_1)

farms_sf_correct <- farms_sf %>%
  mutate(GID_0 = as.character(GID_0)) %>%
  mutate(NAME_0 = as.character(NAME_0)) %>%
  mutate(NAME_1 = as.character(NAME_1)) %>%
  mutate(GID_0 = ifelse(check %in% "CYP:XAD", "CYP", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CYP:XAD", "Cyprus", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "ESP:PRT", "ESP", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "ESP:PRT", "Spain", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "GRC:ALB", "GRC", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "GRC:ALB", "Greece", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "GRC:TUR", "GRC", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "GRC:TUR", "Greece", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "KOR:PRK", "KOR", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "KOR:PRK", "South Korea", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "HKG:CHN", "HKG", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HKG:CHN", "Hong Kong", NAME_0))

farms_sf_correct$check <-  paste(farms_sf_correct$iso3c, farms_sf_correct$GID_0, sep=":")
table(farms_sf_correct$check)

@@

# FAO shrimp production data per country
```{r}

fao <- read.csv(here::here("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv")) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::select(iso3c, fao_tonnes_production)

@@

farms_sf_correct <- farms_sf_correct %>%
  left_join(fao, by="iso3c")

@@

## Egypt is the only one with regional production
farms_rgn_prod <- farms_sf_correct %>%
  dplyr::filter(!is.na(rgn_production))

## calculate farm production
farms_rgn_prod <- farms_rgn_prod %>%
  mutate(prop_production = rgn_production/sum(rgn_production)) %>%
  mutate(regional_prod = fao_tonnes_production*prop_production) %>%
  group_by(NAME_1) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = regional_prod/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry)

@@

## calculate farm production
farms_prod <- farms_sf_correct %>%
  group_by(NAME_0) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = fao_tonnes_production/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry) %>%
  dplyr::filter(iso3c != "EGY") ## filter out EGY since we have regional production estimates

@@

farms_sf_prod <- st_as_sf(farm_production)

## write the production data so that we can use it for the nutrient layer. 
st_write(farms_sf_prod, "marine/marine_fish_general/finfish_nutrient/prod_data_by_cell_iso3c.shp")
```

@@

sum(farms_sf$tonnes_per_farm) # 3901214

finfish_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
plot(log(finfish_raster_check+1))
cellStats(finfish_raster_check, stat = "sum", na.rm = TRUE) # 3901214

finfish_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)
plot(log(finfish_raster+1))

finfish_raster_df <- as.data.frame(finfish_raster, xy=TRUE) %>%
  dplyr::select(x,y, tonnes_production=layer)

cellStats(finfish_raster, stat = "sum") # 3901214
sum(finfish_raster_df$tonnes_production) # 3901214
sum(farms_sf$tonnes_per_farm, na.rm = TRUE) # 3901214
sum(fao$fao_tonnes_production) # 3901214

write_csv(finfish_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/marine_fish_general/marine_fish_general_farm.csv")

>>>>

---
title: "Rasterizing prodution per each region"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/17/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

farms <- read.csv(here::here("marine/crustaceans/crustacean_farms/data/global_crustacean_farm_lat_lon.csv"))

@@

@@

# FAO crustacean production data per country
```{r}

fao <- read.csv(here::here("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans.csv")) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::select(iso3c, fao_tonnes_production)

@@

farms_sf_correct <- farms_sf %>%
  left_join(fao, by="iso3c")

@@

## Vietnam and Philippines have regional production
farms_rgn_prod <- farms_sf_correct %>%
  dplyr::filter(!is.na(prod))

## calculate farm production
farms_rgn_prod <- farms_rgn_prod %>%
  group_by(iso3c) %>%
  mutate(prop_production = prod/sum(prod)) %>%
  mutate(regional_prod = fao_tonnes_production*prop_production) %>%
  group_by(iso3c, sub_rgn) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = regional_prod/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry)

## Check to make sure the totals match
test <- farms_rgn_prod %>%
  dplyr::filter(iso3c == "VNM")

sum(test$tonnes_per_farm)
```

@@

## calculate farm production
farms_prod <- farms_sf_correct %>%
  group_by(NAME_0) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = fao_tonnes_production/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry) %>%
  dplyr::filter(!(iso3c %in% c("VNM", "PHL"))) ## filter out VNM and PHL since we have regional production estimates for those countries

@@

@@

sum(farms_sf$tonnes_per_farm) # 401719.1
sum(fao$fao_tonnes_production) # 401719.1

crustacean_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
plot(log(crustacean_raster_check+1))
cellStats(crustacean_raster_check, stat = "sum", na.rm = TRUE) # 401719.1

crustacean_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)

plot(log(crustacean_raster+1))

crustacean_raster_df <- as.data.frame(crustacean_raster, xy=TRUE) %>%
  dplyr::select(x,y, tonnes_production=layer)

cellStats(crustacean_raster, stat = "sum") # 401719.1
sum(crustacean_raster_df$tonnes_production) # 401719.1
sum(farms_sf$tonnes_per_farm, na.rm = TRUE) # 401719.1
sum(fao$fao_tonnes_production) # 401719.1

write_csv(crustacean_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/crustaceans/crustaceans_farm.csv")

### 2021-03-23 marine_mapping_archive_v1/explore_spatial_allocation.Rmd
Finished reorganizing data

## make a dataset with the country, species type (farm type), tonnage, and # of farms, for those that score data quality %in% c(3,4,5)
sum(all_species_country_tonnes$fao_tonnes_production) # 29623515

>>>>

## make a dataset with the country, species type (farm type), tonnage, and # of farms, for those that score data quality %in% c(3, 4, 5)... review all of these, since many of these are just subsets of total number of farms...

sum(all_species_country_tonnes$fao_tonnes_production) # 29623515

n_farms_final <- n_farms %>%
  filter(data_quality %in% c(3,4,5))
